---
title: "GO-term analysis and plotting for METTL7A"
output: html_notebook
author: 'Christina Fitzsimmons'
date: 'Last updated 2022-05-08'
---

# Introduction
Goal of this library is to plot the GO-terms from Metascape and reviGO collapse
```{r, Load Libraries}
library(tidyverse)
library("RColorBrewer")
library(ggrepel)
library(ggpubr)
library(gprofiler2)

theme_set(theme_bw())

```
# Import GO-term data from METASCAPE
Genes with l2fc 1 and padj < 0.05 from MCF7 and Hut78 RNA-seq data sets were analyzed with [Metascape](https://metascape.org/gp/index.html#/main/step1) and the resulting lists were imported here. 
```{r}

#Import and filter GO-term data
Hut78_down <- read.csv(file = "./2022.05/2022.05_Metascape_Results/2022.05_l2fc1_padj05_Hut78_downregulated/Enrichment_GO/_FINAL_GO.csv", 
                       header = TRUE)

Hut78_up <- read.csv(file = "./2022.05/2022.05_Metascape_Results/2022.05_l2fc1_padj05_Hut78_upregulated/Enrichment_GO/_FINAL_GO.csv",
                     header = TRUE)

MCF7_down <- read.csv(file = "./2022.05/2022.05_Metascape_Results/2022.05_l2fc1_padj05_MCF7_downregulated/Enrichment_GO/_FINAL_GO.csv",
                      header = TRUE)

MCF7_up <- read.csv(file= "./2022.05/2022.05_Metascape_Results/2022.05_l2fc1_padj05_MCF7_upregulated/Enrichment_GO/_FINAL_GO.csv",
                   header = TRUE)
                    
```
The data was filtered for only GO:BiologicalPathway terms and written to a csv file for analysis in [ReviGO](http://revigo.irb.hr/) Revigo will collapse lists of GO-terms by removing redundant ones. 
```{r}

# Filter for GO-biological process
Hut78_down_BP <- Hut78_down %>%
  dplyr::filter(Category == 'GO Biological Processes')
write.csv(Hut78_down_BP, file = "./2022.05/2022.05_Metascape_Results/reviGO_input/Hut78_down_GOBP.csv")

Hut78_up_BP <- Hut78_up %>%
  dplyr::filter(Category == 'GO Biological Processes')
write.csv(Hut78_up_BP, file = "./2022.05/2022.05_Metascape_Results/reviGO_input/Hut78_up_GOBP.csv")

MCF7_down_BP <- MCF7_down %>%
  dplyr::filter(Category == 'GO Biological Processes')
write.csv(MCF7_down_BP, file = "./2022.05/2022.05_Metascape_Results/reviGO_input/MCF7_down_GOBP.csv")

MCF7_up_BP <- MCF7_up %>%
  dplyr::filter(Category == 'GO Biological Processes')
write.csv(MCF7_up_BP, file = "./2022.05/2022.05_Metascape_Results/reviGO_input/MCF7_up_GOBP.csv")

```
# plots of reviGO collapsed data
Revigo collapsed data (collapsed at "small" (0.5) level) were imported to R for plotting. 
```{r}
Hut78_down_revigo <- read_tsv("./2022.05/2022.05_Metascape_Results/reviGO_output/2022.05_revigo_filter05_Hut78_down.tsv")

Hut78_up_revigo <- read_tsv("./2022.05/2022.05_Metascape_Results/reviGO_output/2022.05_revigo_filter05_Hut78_up.tsv")

MCF7_down_revigo <- read_tsv("./2022.05/2022.05_Metascape_Results/reviGO_output/2022.05_revigo_filter05_MCF7_down.tsv")

MCF7_up_revigo <- read_tsv("./2022.05/2022.05_Metascape_Results/reviGO_output/2022.05_revigo_filter05_MCF7_up.tsv")
```

# Revigo collapsed GO-term plots
Hut78 and MCF7
```{r, Hut78 filtering}

# getting things in order
# https://wilkelab.org/SDS375/slides/getting-things-in-order.html#43

Hut78_down_10 <- Hut78_down_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(direction = 'downregulated') %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

Hut78_up_10 <- Hut78_up_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(direction = 'upregulated') %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

Hut78_combined <- bind_rows(Hut78_down_10, Hut78_up_10) %>%
  dplyr::mutate(id = row_number())

```
```{r, MCF7 filtering}

MCF7_down_10 <- MCF7_down_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(direction = 'downregulated') %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

MCF7_up_10 <- MCF7_up_revigo %>%
  dplyr::filter(Eliminated == FALSE) %>%
  dplyr::arrange(Value) %>% 
  dplyr::slice(1:10) %>% # select the first 10 GO-terms
  dplyr::mutate("log10P" = -1* Value) %>%
  dplyr::mutate(direction = 'upregulated') %>%
  dplyr::mutate(TermID = factor(TermID, TermID)) # This ranks the GO-terms in my order, rather than alphabetically. 

MCF7_combined <- bind_rows(MCF7_down_10, MCF7_up_10) 
```

# barplots
```{r}
Hut78_all_plot <- ggplot(Hut78_combined , aes(x = log10P, y =TermID, fill = direction)) +
  geom_col(show.legend = FALSE) +
  facet_grid(direction ~ ., scales = "free") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "Hut78 GO Biological Process",
       subtitle = "reviGO collapse 0.5",
       x = "-Log10P", y = "") +
    scale_y_discrete(limits = rev)
Hut78_all_plot


MCF7_all_plot <- ggplot(MCF7_combined , aes(x = log10P, y = TermID, fill = direction)) +
  geom_col(show.legend = FALSE) +
  facet_grid(direction ~ ., scales = "free") +
  geom_text(aes(label=Name), 
            hjust = 0, x=0, size = 3) +
  labs(title = "MCF7 GO Biological Process",
       subtitle = "reviGO collapse 0.5",
       x = "-Log10P", y = "") +
  scale_y_discrete(limits = rev)
MCF7_all_plot
```
# Exporting bar plots
The resulting plots were exported to pdf format for compilation in illustrator
```{r}
# Exporting the figures
pdf("2022.05.07_Hut78_l2fc1_padj05_revigo07_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
Hut78_all_plot
dev.off()

pdf("2022.05.07_MCF7_l2fc1_padj05_revigo07_barplot.pdf", 
   useDingbats=FALSE,
   width=5, height=4)
MCF7_all_plot
dev.off()
```


